Supplementary Materials

### 2023-08-28 reports/county_figures.Rmd
update lag figures

---
title: "County Figures"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache=TRUE, fig.width=10, fig.height=7)

>>>>

---
title: "County Figures"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message=FALSE, 
                      warning=FALSE,
                      cache=TRUE, 
                      fig.width=10, 
                      fig.height=7)

### 2023-08-28 reports/state_figures.Rmd
update lag figures

---
title: "State Results"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

@@

state_res %>%
  select(-c(date)) %>%
  distinct() %>%
  mutate(interval_length =exp_cases_ub - exp_cases_lb) %>%
  group_by(vlabel) %>%
  summarize(mean_interval_length = mean(interval_length))

@@

>>>>

---
title: "State Results"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
    toc_depth: 4
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output")})
---

@@

state_res %>%
  select(-c(date)) %>%
  distinct() %>%
  mutate(interval_length =exp_cases_ub - exp_cases_lb) %>%
  group_by(vlabel) %>%
  summarize(mean_interval_length = mean(interval_length)) %>%
  arrange(desc(mean_interval_length)) %>%
  kableExtra::kbl()

state_res_differenced <- state_res %>%
  select(-date) %>% distinct() %>%
  mutate(testpos=positive/total) %>%
  group_by(state, vlabel) %>%
  mutate(exp_cases_median_differenced =
           exp_cases_median - lag(exp_cases_median,
                                  n =1,
                                  order_by = biweek),
         infections_differenced = infections - lag(infections,
                                                   n = 1,
                                                   order_by=biweek),
         testpos_differenced = lag(testpos, n=1, order_by=biweek),
         positive_differenced =lag(positive,n=1,order_by=biweek)) %>%
  select(state, name, vlabel, contains('differenced'), biweek) %>%
  filter(!is.na(exp_cases_median_differenced)) %>%
  filter( biweek <=25)

# if lag is NEGATIVE, covidestim LEADS wastewater
# if lag is POSITIVE, covidestim LAGS wastewater
# since by the documentation, 'the lag k value returned by ccf(x, y) estimates
# the correlation between x[t+k] and y[t].'

get_max_lag_by_version <- function(input_df_for_version,
                                   varnames = c("infections_differenced",
                                  "exp_cases_median_differenced"),
                                  comparison = "bias corrected") {

  var1 <- varnames[1]
  var2 <-varnames[2]
  lab <- unique(input_df_for_version$vlabel)

  cross_correlations <- ccf(
    pull(input_df_for_version, var1),
     pull(input_df_for_version, var2), 
                            plot = FALSE,
    lag.max=3)

  ccf_res <- tibble(lag = cross_correlations$lag[,,1], 
                  correlation=cross_correlations$acf[,,1])

  m <- ccf_res %>% filter(correlation == max(correlation))

  tibble(vlabel = unique(input_df_for_version$vlabel),
         fips = unique(input_df_for_version$state),
         max_lag = m$lag,
         max_corr = m$correlation,
         comparison = comparison,
         name = unique(input_df_for_version$name))

cross_corr <- state_res_differenced %>%
  group_by(state, vlabel) %>%
  group_split() %>%
  map_df(get_max_lag_by_version)

states_with_lag <- cross_corr %>%
  filter(max_lag >  0 & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
  # remove those where lag appears to be due to low correlation rather 
  # than true lag
  filter(max_corr > .6) %>% 
  pull(fips) %>%
  unique()

cross_corr_testpos <- state_res_differenced %>%
  filter(biweek >=6 & biweek <= 25) %>%
  # positive tests and positivity are not dependent on vlabel; use this
  # one because it goes to earlier dates
  filter(vlabel=="$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
  group_by(state, vlabel) %>%
  group_split() %>%
  map_df(~get_max_lag_by_version(.x,
    varnames = c("positive_differenced", 
                 "testpos_differenced"))) %>%
  filter(max_lag > 0) %>%
  pull(fips) %>%
  unique()

# length(states_with_lag)
# setdiff(states_with_lag, cross_corr_testpos) %>% length()

```{r, include=FALSE}

state_res_differenced %>%
  filter(state == "AL" & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
  ggplot(aes(x=exp_cases_median_differenced, y = infections_differenced)) +
  geom_point()

state_res %>%
  filter(state == "AL" & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
  ggplot(aes(x=exp_cases_median, y = infections)) +
  geom_point()

@@

## Compare States that Have a Lag

States where the Covidestim estimates are lagged relative to the probabilistic bias counts are `r subset %>% filter(state %in% states_with_lag) %>% pull(name) %>% unique()`.

```{r,  fig.height=11, fig.width=12}

# states_with_lag

lagged_fig_a <- subset %>%
    filter(state %in% states_with_lag & biweek <= 25 & biweek >=6) %>%
   ggplot() +
  geom_ribbon(aes(x = date, 
             ymin = exp_cases_lb,
             ymax = exp_cases_ub,
             fill = 'Probabilistic Bias Analysis'),
             alpha = .8,
             show.legend=FALSE) +
  geom_ribbon(aes(x = date,
                  fill = 'Covidestim',
                  ymin = infections.lo,
                  ymax = infections.hi),
              alpha = .6) +
  # geom_vline(xintercept=end_old_model,
  #            linetype=2,
  #            color="darkred",
  #            linewidth=.5) +
  facet_wrap(~name, ncol=3, scales = "free",
              labeller= labeller(
                vlabel =as_labeller(TeX, default=label_parsed))) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "3 months", 
               date_labels = "%b %Y") +
  theme_c(axis.text.x=element_text(size=7),
          axis.text.y = element_text(size=4.5),
          legend.position = "top",
          legend.text=element_text(size=12),
          strip.text.x=element_text(size=11,
                                    face='plain',
                                    color='white')) +
  theme(plot.title = element_text(face="bold",
                                 hjust = 0, 
                                 size = 13,
                                 margin=margin(0,0,0,0))) +
  scale_fill_manual(values =comp_state_pal, name='')  +
  labs(y = "Estimated Infections",
       title = paste0(
         "(a) Estimated Infections by State in States with a Lag")) +
  guides(color = guide_legend(override.aes = list(linewidth =2),
                              nrow=2,
                              byrow=TRUE))
lagged_fig_a

```{r, fig.height=11, fig.width=12, eval=save_figures}

ggsave(here('figures/intervals-and-covidestim-states-lag.pdf'))

## Test Positivity and Positive Tests in States with a Lag

```{r testpos-lag, fig.height=14, fig.width=12}

############################################
# POSITIVE TESTS AND TEST POSITIVITY
############################################

 subset %>%
  filter(biweek<=25) %>%
  filter(state %in% states_with_lag) %>%
    mutate(testpos = positive/total)

pltlist <- subset %>%
  filter(biweek >= 6 & biweek<=25) %>%
  filter(state %in% states_with_lag) %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  summarize(date=min(date) + days(7)) %>%
  group_by(state) %>%
  mutate(adj = (1/max(testpos)) * max(positive)) %>%
  group_split() %>%
  map(~ {
    adj <-unique(.x$adj)
    .x %>%
      ggplot(aes(x=date))+
    geom_line(aes(y=positive, color = 'Positive Tests'),
              alpha=.8) +
      geom_point(aes(y=positive, color = 'Positive Tests'),
              alpha=.5,
              size=.9) +
    geom_line(aes(y=testpos*adj, color='Test Positivity'),
              alpha=.8
              #,linetype=2
              ) +
      geom_point(aes(y=testpos*adj, color='Test Positivity'),
              alpha=.5,
              size=.9) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Test Positivity"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
                                                    linewidth=c(2,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=10),
          axis.text =element_text(size=9),
          legend.position="none") +
    scale_y_continuous(labels=comma, 
                       sec.axis = sec_axis(
                         ~./adj,
                         name = 'Test Positivity')) +
    labs(color ='',
         y="Positive Tests",
         x= "Date") +
      facet_wrap(~name) +
      scale_x_date(date_labels="%b %Y",
                   date_breaks = "3 months")

  })

legend <- get_legend(
  pltlist[[1]] +
    theme(legend.position = "top",
          legend.text=element_text(size=13, margin=margin(6,3,0,0)))
)

title_gg <- ggplot() +
  labs(
    #title = "(b) Comparing Trends in Positive Tests and Test Positivity"
    title = "") +
  theme(plot.title=element_text(face="bold",
                                hjust = 0,
                                size = 13))

plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)

lagged_fig_b <- plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)

# lagged_fig_b <- plot_grid( legend, plt, rel_heights=c(.05,.95), ncol=1)

lagged_fig_b

### States Where Covidestim Estimates Appear to be Lagged Relative to Probabilistic Bias Counts
```{r, eval=save_figures, fig.height=14.8, fig.width=12}

lag_title <- ggplot() +
  labs(title = "States Where Covidestim Estimates Appear to be Lagged\n Relative to Probabilistic Bias Counts") +
  theme(plot.title=element_text(face="bold", hjust = .5, size=18))

plot_grid(# lag_title, 
          lagged_fig_a + labs(title=""), 
          lagged_fig_b,
          labels=c(
           # "",
            "(a) Estimated Infections by State in States with a Lag",
            "(b) Comparing Trends in Positive Tests and Test Positivity in States with a Lag"),
        #  rel_heights = c(.04, .47,.49),
          label_x=0,
         # vjust=c(1.5,1.5,.1),
        vjust=c(1.5,.1),
          hjust=-.05,
          ncol=1)

ggsave(here('figures/testpos-lag.pdf'))

########################################################
# PERCENT CHANGE IN POSITIVE TESTS AND TEST POSITIVITY
########################################################

 subset %>%
  filter(biweek<=25) %>%
  filter(state %in% states_with_lag) %>%
    mutate(testpos = positive/total)

pltlist <- subset %>%
  filter(biweek >= 6 & biweek<=25) %>%
  filter(state %in% states_with_lag) %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  # put date in midpoint
  summarize(date=min(date) + days(7)) %>%
  group_by(state) %>%
  mutate(pct_change_testpos = (testpos-lag(testpos, n=1, order_by=biweek)) /
                                 lag(testpos, n=1,
                                     order_by=biweek),
         pct_change_positive = (positive - lag(positive,
                                               n=1,
                                               order_by=biweek)) / 
                                 lag(positive,
                                               n=1,
                                               order_by=biweek)) %>%
 #  mutate(adj = (1/max(testpos)) * max(positive)) %>%
  group_split() %>%
  map(~ {
   # adj <-unique(.x$adj)
    .x %>%
      ggplot(aes(x=date))+
    geom_line(aes(y=pct_change_positive, color = 'Positive Tests'),
              alpha=.8) +
      geom_point(aes(y=pct_change_positive, color = 'Positive Tests'),
              alpha=.5,
              size=.9) +
    geom_line(aes(y=pct_change_testpos, color='Test Positivity'),
              alpha=.8
              #,linetype=2
              ) +
      geom_point(aes(y=pct_change_testpos, color='Test Positivity'),
              alpha=.5,
              size=.9) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Test Positivity"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
                                                    linewidth=c(2,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=10),
          axis.text =element_text(size=9),
          legend.position="none") +
    # scale_y_continuous(labels=comma, 
    #                    sec.axis = sec_axis(
    #                      ~./adj,
    #                      name = 'Test Positivity')) +
    labs(color ='',
         y="Positive Tests",
         x= "Date") +
      facet_wrap(~name) +
      scale_x_date(date_labels="%b %Y")

  })

legend <- get_legend(
  pltlist[[1]] +
    theme(legend.position = "top",
          legend.text=element_text(size=16))
)

title_gg <- ggplot() + 
  labs(title = "Comparing Trends in Positive Tests and Test Positivity") + 
  theme(plot.title=element_text(face="bold",
                                hjust = .5, 
                                size = 18,
                                margin =margin(5,0,2,0)))

plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)

plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)

########################################################
# PERCENT CHANGE IN POSITIVE TESTS AND TOTAL TESTS
########################################################

pltlist <- subset %>%
  filter(biweek >= 6 & biweek<=25) %>%
  filter(state %in% states_with_lag) %>%
    mutate(testpos = positive/total) %>%
  group_by(across(-date)) %>%
  # put date in midpoint
  summarize(date=min(date) + days(7)) %>%
  group_by(state) %>%
  mutate(pct_change_testpos = (total-lag(total, n=1, order_by=biweek)) /
                                 lag(total, n=1,
                                     order_by=biweek),
         pct_change_positive = (positive - lag(positive,
                                               n=1,
                                               order_by=biweek)) / 
                                 lag(positive,
                                               n=1,
                                               order_by=biweek)) %>%
 #  mutate(adj = (1/max(testpos)) * max(positive)) %>%
  group_split() %>%
  map(~ {
   # adj <-unique(.x$adj)
    .x %>%
      ggplot(aes(x=date))+
    geom_line(aes(y=pct_change_positive, color = 'Positive Tests'),
              alpha=.8) +
      geom_point(aes(y=pct_change_positive, color = 'Positive Tests'),
              alpha=.5,
              size=.9) +
    geom_line(aes(y=pct_change_testpos, color='Total Tests'),
              alpha=.8
              #,linetype=2
              ) +
      geom_point(aes(y=pct_change_testpos, color='Total Tests'),
              alpha=.5,
              size=.9) +
    #scale_linetype_discrete(breaks = c(1,1,2)) +
    scale_color_manual(values=c(
                                "Positive Tests"= "#279143",
                                "Total Tests"= '#E38E4F')) +
    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
                                                    linewidth=c(2,2)))) +
    theme_c() +
    theme(axis.title = element_text(size=10),
          axis.text =element_text(size=9),
          legend.position="none") +
    # scale_y_continuous(labels=comma, 
    #                    sec.axis = sec_axis(
    #                      ~./adj,
    #                      name = 'Test Positivity')) +
    labs(color ='',
         y="Total Tests",
         x= "Date") +
      facet_wrap(~name) +
      scale_x_date(date_labels="%b %Y")

  })

legend <- get_legend(
  pltlist[[1]] +
    theme(legend.position = "top",
          legend.text=element_text(size=16))
)

title_gg <- ggplot() + 
  labs(title = "Comparing Trends in Positive Tests and Total Tests") + 
  theme(plot.title=element_text(face="bold",
                                hjust = .5, 
                                size = 18,
                                margin =margin(5,0,2,0)))

plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)

plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)

### 2023-08-28 writing/supplement.Rmd
update lag figures

---
title: "Supplementary Materials"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: true
    number_sections: false

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
library(knitr)
library(here)

@@

```{r ctis-by-state, fig.cap ="Since $\\beta$ represents the ratio of $\\Pr(\\text{test}_+|\\text{untested}, S_0)$ to $\\Pr(\\text{test}_+|\\text{tested})$, we estimate $\\beta$ from the COVID-19 Trends and Impact Survey as the ratio of the screening test positivity over the overall test positivity. Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state. Most have a mean near $20%$."}

@@

```{r testrate-low-summer, fig.cap="Testing rate, calculated as the etotal number of tests for a 2-week interval in a state over the census population in that state, over time. The median across states for each two-week interval is shown in red. In relation to Figure 1 in the main text, where we see that the period with the highest ratio of estimated to observed infections is June 18, 2021 through July 1, 2021, we see here that testing rates across states are at a minimum during this time period."}

@@

include_graphics(here('figures/intervals-and-covidestim-states-lag.pdf'))

@@

@@

```{r michigan-variant, fig.cap = "Simulation intervals for Michigan along with the variant proportions for variants designated as *Variant of Interest* or *Variant of Concern* by the World Health Organization. These variant proportions are not specific to Michigan; the estimates are for the United States. The rise in Omicron corresponds to the dramatic increase in both the probabilistic bias intervals and Covidestim estimates in December of 2021 through January of 2022. We also see the increase in the Delta variant that precedes the Delta wave in the summer of 2021."}

>>>>

---
title: "Supplementary Materials"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: true
    number_sections: false
header-includes:
- \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.pos="H",out.extra = "")
library(knitr)
library(here)

@@

```{r ctis-by-state, fig.cap ="Since $\\beta$ represents the ratio of $\\Pr(\\text{test}_+|\\text{untested}, S_0)$ to $\\Pr(\\text{test}_+|\\text{tested})$, we estimate $\\beta$ from the COVID-19 Trends and Impact Survey as the ratio of the screening test positivity over the overall test positivity. Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state. Most have a mean near 0.20."}

@@

```{r testrate-low-summer, fig.cap="Testing rate, calculated as the total number of tests for a 2-week interval in a state over the census population in that state, over time. The median across states for each two-week interval is shown in red. In relation to Figure 1 in the main text, where we see that the period with the highest ratio of estimated to observed infections is June 18, 2021 through July 1, 2021, we see here that testing rates across states are at a minimum during this time period."}

@@

# include_graphics(here('figures/intervals-and-covidestim-states-lag.pdf'))

@@

```{r lagged-figs, fig.cap="States where the Covidestim infection estimates are lagged relative to the probabilistic bias estimates, that is, they are most highly correlated when the Covidestim infections are lagged by one or more biweeks. (a) Simulation intervals demonstrating the lag seen around the delta peak in summer 2021. (b) Test positivity and positive tests for each state. Increases in test positivity tend to precede increases in positive tests, and these increases in test positivity result in increases in the estimates probduced by the probabilistic bias analysis seen in part (a)."}

@@

```{r michigan-variant, fig.cap = "Simulation intervals for Michigan along with the variant proportions for variants designated as \\emph{Variant of Interest} or \\emph{Variant of Concern} by the World Health Organization. These variant proportions are not specific to Michigan; the estimates are for the United States. The rise in Omicron corresponds to the dramatic increase in both the probabilistic bias intervals and Covidestim estimates in December of 2021 through January of 2022. We also see the increase in the Delta variant that precedes the Delta wave in the summer of 2021."}
